---
title: "p8105_hw6_hp2661"
author: "Huizhong Peng"
date: "2024-12-02"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(ggplot2)
library(modelr)
library(mgcv)

options(scipen = 999)

knitr::opts_chunk$set(
        collapse = TRUE, 
        message = FALSE, 
        warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
  )

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}

quantity = function(model) {
  r_sq = broom::glance(model)$r.squared
  beta_0 = broom::tidy(model) |> filter(term == "(Intercept)") |> pull(estimate)
  beta_1 = broom::tidy(model) |> filter(term == "tmin") |> pull(estimate)
  log_beta = log(abs(beta_0*beta_1))
  
  tibble(r_squared = r_sq, log_beta)
}

boot_straps = 
  tibble(strap_number = 1:5000) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(df = weather_df))
  )

bootstrap_results = 
  boot_straps |> 
  mutate(
    models = map(strap_sample, \(df) lm(tmax ~ tmin, data = df)),
    results = map(models, quantity)
  ) |> 
  select(-strap_sample, -models) |> 
  unnest(results) 

# plot
bootstrap_results |> 
  ggplot(aes(x = r_squared)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Distribution of r^2", x = "r^2", y = "Frequency")

bootstrap_results |> 
  ggplot(aes(x = log_beta)) +
  geom_histogram(fill = "pink", color = "black") +
  labs(title = "Distribution of log beta", x = "log beta", y = "Frequency")
```

These quantities are normal distributed.

### Problem 2

Baltimore:

```{r}
url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/refs/heads/master/homicide-data.csv"
raw_dat = read_csv(url, na = c("Unknown", ""))

write.csv(raw_dat, "data/homicide_raw_data.csv", row.names = FALSE)

# data cleaning
dat = raw_dat |> 
  janitor::clean_names() |> 
  mutate(
    city_state = paste(city, state, sep = ", "),
    resolved = ifelse(disposition == "Closed by arrest", 1, 0),
    victim_age = as.numeric(victim_age)
  ) |> 
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) |> 
  filter(victim_race %in% c("White", "Black"))

bal = dat |> 
  filter(city_state == "Baltimore, MD") |> 
  select(uid, resolved, victim_age, victim_sex, victim_race)

# glm
bal_fit_log = bal |> 
  glm(resolved ~ victim_age + victim_sex + victim_race, 
      data = _, family = binomial())

bal_fit_log |> 
  broom::tidy() |> 
  knitr::kable(digits = 3)

bal_sex_OR = 
  bal_fit_log |>
  broom::tidy(conf.int = TRUE, exponentiate = TRUE) |> 
  select(term, OR = estimate, std = std.error, 
         ci_low = conf.low, ci_high = conf.high) |> 
  filter(term == "victim_sexMale") |> 
  knitr::kable(digits = 3)

bal_sex_OR
```

All cities:

```{r}
fit_log = function(data){
  model = glm(resolved ~ victim_age + victim_sex + victim_race, 
              data = data, family = binomial())
  
  model |> 
    broom::tidy(conf.int = TRUE, exponentiate = TRUE) |> 
    select(term, OR = estimate, std = std.error, 
           ci_low = conf.low, ci_high = conf.high) |> 
    filter(term == "victim_sexMale")
}

dat_nest = dat |> 
  group_by(city_state) |> 
  nest() 

all_sex_OR = 
  dat_nest |> 
  mutate(
    model = map(data, fit_log)
  ) |> 
  unnest(model) |> 
  select(city_state, OR, std, ci_low, ci_high) |> 
  tibble()

all_sex_OR |> 
  knitr::kable(digits = 3)
```

Plot: 

```{r}
all_sex_OR |> 
  mutate(
    city_state = fct_reorder(city_state, OR)
  ) |> 
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0.5) +
  labs(
    title = "Adjusted Odds Ratios for Solving Homicides by City",
    x = "City",
    y = "Odds Ratio (Male vs Female)"
  ) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1)
  )
```

Adjusted odds Ratios in most cities are less than 1, showing that there are more female victims than male victims keeping all other variables fixed.

New York City has the lowest adjusted OR and Albuquerque City has the highest adjusted OR.

### Problem 3

```{r}
bwt_df = 
  read_csv("data/birthweight.csv", 
           col_types = cols(
             babysex = "c", frace = "c", malform = "c", mrace = "c"))

skimr::skim(bwt_df)

model = lm(bwt ~ bhead + blength + babysex + delwt + fincome + frace + gaweeks + malform + mrace + ppbmi + ppwt + smoken + wtgain, 
         data = bwt_df)
summary(model)
```

Remain significant variables (e.g., `bhead`, `blength`, `babysex`) in final regression model.

```{r}
final_model = lm(bwt ~ bhead + blength + babysex + delwt + gaweeks + mrace + ppbmi + smoken,
              data = bwt_df)
summary(final_model)

bwt_df = bwt_df |> 
  add_predictions(final_model) |> 
  add_residuals(final_model)

# plot
bwt_df |> 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Residuals vs Fitted Values",
    x = "Fitted Values",
    y = "Residuals"
  )
```

Compare this model to other models:

```{r}
model1 = lm(bwt ~ blength + gaweeks,
            data = bwt_df)
summary(model1)
model2 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex,
            data = bwt_df)
summary(model2)

cv_df = 
  crossv_mc(bwt_df, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_df = 
  cv_df |> 
  mutate(
    mod_f = map(train, \(df) final_model),
    mod_1 = map(train, \(df) model1),
    mod_2 = map(train, \(df) model2)) |> 
  mutate(
    rmse_my = map2_dbl(mod_f, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_main = map2_dbl(mod_1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_interaction = map2_dbl(mod_2, test, \(mod, df) rmse(model = mod, data = df)))

cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

The model containing `bhead`, `blength`, `babysex`, `delwt`, `gaweeks`, `mrace`, `ppbmi` and `smoken` has the lowest rmse. 
